---
title: "Results Stat Analysis"
author: "Lucy and Sammie"
date: "25/11/2021"
output: pdf_document
---

Load libraries:
```{r, include=FALSE}
library(ggplot2)
library(popbio)
library(dplyr)
library(patchwork)
library(DHARMa)
library(mgcv)
library(fitdistrplus)
library(goft)
library(gamlss)
library(viridis)
library(tidyverse)
library(patchwork)
```

Loading colour data and cleaning
```{r}
  
ct_data <- read.csv("colour_data_DS2021.csv") %>%
  mutate(id = as.factor(id)) %>%
  dplyr::select("id", "change_g7", "species", "treatment") %>%
  na.omit(id, change_g7, species, treatment, int_mean, event) %>%
  mutate(treatment = fct_relevel(treatment, "control", "low", "mid", "high")) %>%
   mutate(species = fct_relevel(species, "cham", "calli","chillensis"))
  

```

filtering for species and cleaning more 
```{r}
cham_ctdata <- ct_data %>%
  mutate(treatment = fct_relevel(treatment, "control", "low", "mid", "high")) %>%
  dplyr::filter(species == "cham") 

ca_ctdata <- ct_data %>%
  mutate(treatment = fct_relevel(treatment, "control", "low", "mid", "high")) %>%
  dplyr::filter(species == "calli")

chil_ctdata <- ct_data %>%
  mutate(treatment = fct_relevel(treatment, "control", "low", "mid", "high")) %>%
  dplyr::filter(species == "chillensis")
```


Normality test for colour data
```{r}
shapiro.test(ct_data$change_g7)
```
The colour data is not normal 


Collen & Frey Graph of colour data
```{r}
descdist(ct_data$change_g7)
```
From the Collen and Frey graph we will try gamma, weibull, lognormal, and a beta 

Looking at other distribution fits
```{r}
fitDist(change_g7, data = ct_data, type = "realAll", try.gamlss = T)
```
Initially suggestd a reverse gumbel and now a skewed normal type 2 distribution is suggested so we will test both for the lowest AIC.


Look at histograms to find distribution: 
```{r}
cmGA <- histDist(ct_data$change_g7, "GA", density = T, main = "Gamma")
cmlN <- histDist(ct_data$change_g7, "LOGNO", density = T, main = "lNorm")
cmWE <- histDist(ct_data$change_g7, "WEI", density = T, main = "Weibull")
cmB <- histDist(ct_data$change_g7, "BE", density = T, main = "Beta")
cRG <- histDist(ct_data$change_g7, "RG", density = T, main = "Reverse Gumbel")
cNO <- histDist(ct_data$change_g7, "NO", density = T, main = "Normal")
cSN2<- histDist(ct_data$change_g7, "SN2", density = T, main = "Skew Normal 2")
```
Found out that some of the models do not work due to negative values possibly and will continue GAIC using reverse gumbel vs. Skewed noraml 2, and normal
AIC 
```{r}
GAIC(cRG,cNO,cSN2)
```

Lowest AIC is Skewed normal 2 but is within less than 2 points of Reverse gumbel AIC, and Reverse gumbel has a lower df, so we will begin modelling with Reverse Gumbel!

```{r}
cmRG <- histDist(ct_data$change_g7, "RG", density = T, main = "Reverse Gumbel")
```

Making a model!
```{r}
mod_RG <- gamlss(change_g7 ~ species + treatment + random(id), family = RG(), data = ct_data)

summary(mod_RG)
```

Comparing the change in intensity means for significance (by treatment for species)
```{r}

library(FSA) 
library(fGarch)
library(LambertW)

kruskal.test(change_g7 ~ treatment, data = ct_data)

kruskal.test(change_g7 ~ treatment, data = cham_ctdata)
dunnTest(change_g7 ~ treatment, data = cham_ctdata)

kruskal.test(change_g7 ~ treatment, data = ca_ctdata)
dunnTest(change_g7 ~ treatment, data = ca_ctdata)

kruskal.test(change_g7 ~ treatment, data = chil_ctdata)
dunnTest(change_g7 ~ treatment, data = chil_ctdata)
```

Siginificance found! Found for C. Chilensis and C. tuberculosum
For C. tub it is significant between control-high (p=0.011) and ALMOST for control-mid (p=0.05457)
For C. chil it is significant between control-high(p=), high-low(p=0.0036), control-mid(p=0.0175), low-mid (p=0.3274)


Making boxplots for change in mean intensity by treatment for each species
```{r}
cham_ctgplot <- ggplot(cham_ctdata, aes(x= treatment, y = change_g7, fill = treatment)) +
  geom_boxplot() + 
  guides(fill = FALSE) +
  ylab("Change in mean intensity") +
  xlab("H202 Concentration") +
  scale_x_discrete(labels = c('Control','Low','Mid','High')) + 
   ylim(-15,30) +
  ggtitle("A) C. tumidum") +
  scale_fill_manual(values = c('#3C507BFF', '#18DBC5FF', '#F7C13AFF', '#B11901FF'), labels = c("Control", "Low", "Mid", "High")) +
  theme_classic()

ca_ctgplot <- ggplot(ca_ctdata, aes(x= treatment, y = change_g7, fill = treatment)) +
  geom_boxplot() + 
  guides(fill = FALSE) +
  ylab("") +
 xlab("H202 Concentration") +
  labs(fill = "H202 Concentration") +
  scale_x_discrete(labels = c('Control','Low','Mid','High')) + 
   ylim(-15,30) +
  ggtitle("B) C. tuberculosum") +
  scale_fill_manual(values = c('#3C507BFF', '#18DBC5FF', '#F7C13AFF', '#B11901FF'), labels = c("Control", "Low", "Mid", "High")) +
  theme_classic()

chil_ctgplot <- ggplot(chil_ctdata, aes(x= treatment, y = change_g7, fill = treatment)) +
  geom_boxplot() + 
  ylab("") +
 xlab("H202 Concentration") +
  labs(fill = "H202 Concentration") +
  scale_x_discrete(labels = c('Control','Low','Mid','High')) + 
 ylim(-15,30) +
  ggtitle("C) C. chilensis") +
  scale_fill_manual(values = c('#3C507BFF', '#18DBC5FF', '#F7C13AFF', '#B11901FF'), labels = c("Control", "Low", "Mid", "High")) +
  theme_classic()
```


Saving and viewing boxplot
```{r}
colour_fig1 <- cham_ctgplot + ca_ctgplot + chil_ctgplot 
show(colour_fig1)


ggsave(colour_fig1, filename = "colour_figure1.jpg", width = 13, height = 10)
```

Reformating boxplot into a different aestatic
```{r}
species_names <- list(
  'cham'="A) C. tumidum",
  'calli'="B) C. tuberculosum",
  'chillensis'="C) C. chilensis"
)

species_labeller <- function(variable,value){
  return(species_names[value])
}


test_ctgplot <- ggplot(ct_data, aes(x= treatment, y = change_g7, fill = treatment)) +
  geom_boxplot() + 
  ylab("Change in mean intensity") +
  xlab("H2O2 Concentration") + 
  labs(fill = "H2O2 Concentration") +
  facet_grid(. ~ species, labeller=species_labeller) +
  scale_x_discrete(labels = c('Control','Low','Mid','High')) +
  scale_fill_manual(values = c('#3C507BFF', '#18DBC5FF', '#F7C13AFF', '#B11901FF'), labels = c("Control", "Low", "Mid", "High")) +
theme (panel.background = element_rect(fill = NA, color = "black"),strip.text.x = element_text(size = 20, face = "italic"),legend.key.size = unit(2, 'cm'), axis.text.x = element_text(size = 15, angle=90, hjust=0.5), axis.title.x = element_text(size = 25), axis.title.y = element_text(size = 25), legend.text = element_text(size = 14), legend.title = element_text(size = 20), axis.text.y = element_text(size = 20))
```

Viewing and saving recently made plot
```{r}
test_ctgplot

ggsave(test_ctgplot, filename = "colour_figure1facet.jpg", width = 15, height = 10)

```



Subsetting data to compare significance between species 
```{r}
high_ctdata <- ct_data %>%
  dplyr::filter(treatment == "high")

mid_ctdata <- ct_data %>%
  dplyr::filter(treatment == "mid") 

low_ctdata <- ct_data %>%
  dplyr::filter(treatment == "low") 

control_ctdata <- ct_data %>%
  dplyr::filter(treatment == "control") 

```


Testing between species for change in mean intensity significance
```{r}
kruskal.test(change_g7 ~ species, data = ct_data)
dunnTest(change_g7 ~ species, data = ct_data)
  
kruskal.test(change_g7 ~ species, data = control_ctdata)
dunnTest(change_g7 ~ species, data = control_ctdata)

kruskal.test(change_g7 ~ species, data = low_ctdata)
dunnTest(change_g7 ~ species, data = low_ctdata)

kruskal.test(change_g7 ~ species, data = mid_ctdata)
dunnTest(change_g7 ~ species, data = mid_ctdata)

kruskal.test(change_g7 ~ species, data = high_ctdata)
dunnTest(change_g7 ~ species, data = high_ctdata)
```
All species comparisons at each treatment level are significant when compared to chamberlainium

Making boxplot of species by treatment 
```{r}
treatment_names <- list(
  'control'="A) Control",
  'low'="B) Low",
  'mid'="C) Mid",
  'high' ="D) High"
)

treatment_labeller <- function(variable,value){
  return(treatment_names[value])
}


tester_colour_plot <- ggplot(ct_data, aes(x= species, y= change_g7, fill= species)) +
    ylab("Change in mean intensity") +
  xlab("Species") + 
  labs(fill = "Species") + 
  facet_grid(. ~ treatment, labeller = treatment_labeller) +
  scale_x_discrete(labels = c('C. tumidum','C. tuberculosum','C. chilensis')) +
  geom_boxplot() +
  ylab("Change in mean intensity") +
  xlab("Species") + 
  labs(fill = "Species") + 
  scale_fill_manual(values = c('dark magenta', '#E7298A', 'darkorange1'), labels = c('C. tumidum','C. tuberculosum','C. chilensis')) +
theme (panel.background = element_rect(fill = NA, color = "black"),strip.text.x = element_text(size = 20), legend.key.size = unit(2, 'cm'), axis.text.x = element_text(size = 20, angle=90, hjust=0.5, face = "italic"), axis.title.x = element_text(size = 30), axis.title.y = element_text(size = 30), legend.text = element_text(size = 14, face = "italic"), legend.title = element_text(size = 20), axis.text.y = element_text(size = 20), )

```

Viewing and saving previously made plot 
```{r}
tester_colour_plot

ggsave(tester_colour_plot, filename = "colour_figure2.png", height = 10, width = 15)
```











