---
title: "Results Stat Analysis"
author: "Lucy"
date: "25/11/2021"
output: pdf_document
---

Load libraries:
```{r, include=FALSE}
library(ggplot2)
library(popbio)
library(dplyr)
library(patchwork)
library(DHARMa)
library(mgcv)
library(fitdistrplus)
library(goft)
library(gamlss)
library(viridis)
library(tidyverse)
library(rstatix)
```

Load in data and rename PAM column:
```{r}
data <- read.csv("test_data_NOV25.csv")
data <- data %>%
rename("Y_II_PAM" = "Y_II..PAM.") %>%
  mutate(treatment = fct_relevel(treatment, "control", "low", "mid", "high")) %>%
  mutate(species = fct_relevel(species, "cham", "calli ", "chillensis"))
```

Setting up data with PAM columns only:
```{r}
data_to_model <- data %>%
  unite(id, species, species_card, sep = '_', remove = F) %>%
  dplyr::select("date","yellow_tub","species_card","species","number","treatment","name","event","Y_II_PAM","id") %>%
  mutate(id = as.factor(id))
```

re-ordering event for graphing later:
```{r}
data_to_model <- data_to_model %>% mutate(event = fct_relevel(event, "pre_stress", "post_stress", "post_stress_1", "post_stress_2", "post_stress_3", "post_stress_4", "post_stress_5", "post_stress_6","post_stress_7"))
```

testing for normality:
```{r}
shapiro.test(data_to_model$Y_II_PAM)

#not normal P=3.09e-08
```

Cullen & Frey graph:
```{r}
descdist(data_to_model$Y_II_PAM)
```
Looks like it might be a gamma/lognormal/weibull

Look at histograms to find distribution:
```{r}
mGA <- histDist(data_to_model$Y_II_PAM, "GA", density = T, main = "Gamma")
mlN <- histDist(data_to_model$Y_II_PAM, "LOGNO", density = T, main = "lNorm")
mWE <- histDist(data_to_model$Y_II_PAM, "WEI", density = T, main = "Weibull")
mNO <- histDist(data_to_model$Y_II_PAM, "NO", density = T, main = "Normal")
```
```{r}
GAIC(mGA, mlN, mWE, mNO, mSEP2)
#these give us negative values, most negative is best fit - mSEP2, but WEI is really close (we chose to move forward with WEI)
```


```{r}
fitDist(Y_II_PAM, data = data_to_model, type = "realAll", try.gamlss = T)
```

Trying skewed exponential power type 2:
```{r}
mSEP2 <- histDist(data_to_model$Y_II_PAM, "SEP2", density = T, main = "SEP2")
```
Decided to move forward exploring SEP2 and WEI distributions:
```{r}
#mod_SEP2 <- gamlss(Y_II_PAM ~ species + treatment*event + random(id), family = SEP2(), data = data_to_model, control = gamlss.control(n.cyc = 200))

#summary(mod_SEP2)
```
Model using WEI-
```{r}
#mod_WEI <- gamlss(Y_II_PAM ~ species + treatment*event + random(id), family = WEI(), data = data_to_model, method = RS(), control = gamlss.control())

#summary(mod_WEI)
```
```{r}
#seeing if stepGAIC makes a simpler model
#step_mod <- stepGAIC(mod_WEI, direction = "backward", trace = F)
#formula(step_mod)
#step_mod$anova
#summary(step_mod)
```

modelling stripped down data -just 3 days: pre, post1, and post7
```{r}
simpler_data <- data_to_model %>%
  filter(event == "pre_stress" | event == "post_stress_1" | event == "post_stress_7")

simpler_mod_WEI <- gamlss(Y_II_PAM ~ species + treatment*event + random(id), family = WEI(), data = simpler_data, method = RS(), control = gamlss.control())

summary(simpler_mod_WEI)
```
#intercept is = Calliarthron, Control, Pre-stress

```{r}
#seeing if stepGAIC makes a simpler model for this one
step_simpler_mod <- stepGAIC(simpler_mod_WEI, direction = "backward", trace = F)
formula(step_simpler_mod)
step_simpler_mod$anova
summary(step_simpler_mod)
#ASK WHAT THE BENEFIT OF DOING THIS???**** why would we use full one vs reversed one??
```
#For this set of results we are moving forward with the 3-day model including species 
 

box plot for chamberlanium--
```{r}
ch_f <- filter(data_to_model, species == "cham") %>%
  na.omit()

ch_f$event = as.factor(ch_f$event)

ch_f <- ch_f %>% mutate(treatment = fct_relevel(treatment, "control", "low", "mid", "high"))

head(ch_f)

cham_plot<- ggplot(ch_f, aes(x= event, y= Y_II_PAM,fill = treatment)) +
  geom_boxplot()
  
cham_plot
```
box plot for calliarthron--
```{r}
calli_f <- filter(data_to_model, species == "calli ") %>%
  na.omit()

calli_f$event = as.factor(calli_f$event)

calli_f <- calli_f %>% mutate(treatment = fct_relevel(treatment, "control", "low", "mid", "high"))

head(calli_f)

calli_plot<- ggplot(calli_f, aes(x= event, y= Y_II_PAM,fill = treatment)) +
  geom_boxplot()
  
calli_plot
```
box plot for chilensis--
```{r}
chil_f <- filter(data_to_model, species == "chillensis") %>%
  na.omit()

chil_f$event = as.factor(chil_f$event)

chil_f <- chil_f %>% mutate(treatment = fct_relevel(treatment, "control", "low", "mid", "high"))

head(chil_f)

chil_plot<- ggplot(chil_f, aes(x= event, y= Y_II_PAM,fill = treatment)) +
  geom_boxplot()
  
chil_plot
```

trying to make box+whisker?
```{r}
library(plotrix)
library(tidyverse)

#Chamberlanium
ch_simple <- filter(ch_f, event == "pre_stress" | event == "post_stress_1" | event == "post_stress_2" | event == "post_stress_3" | event == "post_stress_4" | event == "post_stress_5" | event == "post_stress_6" | event == "post_stress_7")

ch_for_plot <- ch_simple %>%
group_by(treatment, event) %>%
mutate(mean = mean(Y_II_PAM), SE = std.error(Y_II_PAM))
  

chamberlanium_plot <- ggplot(data = ch_for_plot, aes(x = event, y = mean, group = treatment, color = treatment)) +
  geom_point(aes(shape=treatment, color=treatment), size = 2.5) +
  ylim(0,0.4) +
  scale_color_manual(values=c('#3C507BFF', '#18DBC5FF', '#F7C13AFF', '#B11901FF')) +
  scale_shape_manual(values=c(1,16,17,15)) +
  scale_fill_discrete(name = "H202 Concentration", labels = c("Control", "Low", "Mid", "High")) +
  ggtitle("A) C. tumidum") +
  xlab("Monitoring Day") +
  ylab("YII") +
  geom_linerange(aes(ymin = mean - SE, ymax = mean + SE)) +
  geom_line() +
  theme_gray() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(legend.position = "none") +
  scale_x_discrete(labels=c("Pre-stress", "1", "2", "3", "4", "5", "6", "7")) 
  

chamberlanium_plot

```

```{r}
#Calliarthron
calli_simple <- filter(calli_f, event == "pre_stress" | event == "post_stress_1" | event == "post_stress_2" | event == "post_stress_3" | event == "post_stress_4" | event == "post_stress_5" | event == "post_stress_6" | event == "post_stress_7")

calli_for_plot <- calli_simple %>%
group_by(treatment, event) %>%
mutate(mean = mean(Y_II_PAM), SE = std.error(Y_II_PAM))
  

calliarthron_plot <- ggplot(data = calli_for_plot, aes(x = event, y = mean, group = treatment, color = treatment)) +
  geom_point(aes(shape=treatment, color=treatment), size = 2.5) +
  ylim(0,0.4) +
  scale_color_manual(values=c('#3C507BFF', '#18DBC5FF', '#F7C13AFF', '#B11901FF')) +
  scale_shape_manual(values=c(1,16,17,15)) +
  scale_fill_discrete(name = "H202 Concentration", labels = c("Control", "Low", "Mid", "High")) +
  ggtitle("B) C. tuberculosum") +
  xlab("Monitoring Day") +
  ylab("YII") +
  geom_linerange(aes(ymin = mean - SE, ymax = mean + SE)) +
  geom_line() +
  theme_gray() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(legend.position = "none") +
  scale_x_discrete(labels=c("Pre-stress", "1", "2", "3", "4", "5", "6", "7"))   

calliarthron_plot
```

```{r}
#Chilensis
chil_simple <- filter(chil_f, event == "pre_stress" | event == "post_stress_1" | event == "post_stress_2" | event == "post_stress_3" | event == "post_stress_4" | event == "post_stress_5" | event == "post_stress_6" | event == "post_stress_7")

chil_for_plot <- chil_simple %>%
group_by(treatment, event) %>%
mutate(mean = mean(Y_II_PAM), SE = std.error(Y_II_PAM))
  

chilensis_plot <- ggplot(data = chil_for_plot, aes(x = event, y = mean, group = treatment, color = treatment)) +
  geom_point(aes(shape=treatment, color=treatment), size = 2.5) +
  ylim(0,0.4) +
  scale_color_manual(values=c('#3C507BFF', '#18DBC5FF', '#F7C13AFF', '#B11901FF')) +
  scale_shape_manual(values=c(1,16,17,15)) +
  scale_fill_discrete(name = "H202 Concentration", labels = c("Control", "Low", "Mid", "High")) +
  ggtitle("C) C. chilensis") +
  xlab("Monitoring Day") +
  ylab("YII") +
  theme(legend.title = "H202 Concentration") +
  geom_linerange(aes(ymin = mean - SE, ymax = mean + SE)) +
  geom_line() +
  theme_gray() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  scale_x_discrete(labels=c("Pre-stress", "1", "2", "3", "4", "5", "6", "7")) 

chilensis_plot
```

```{r}

chamberlanium_plot + calliarthron_plot + chilensis_plot
```
Making boxplots for each treatment, comparing species:
```{r}

proper_names <- c(cham = "C. tumidum", "calli " = "C. tuberculosum", chillensis = "C. chilensis")

#high
high_data <- data_to_model %>%
  filter(treatment == "high") %>%
  filter(event == "pre_stress" | event == "post_stress_1" | event == "post_stress_7") %>%
  mutate(species = fct_relevel(species, "cham", "calli ", "chillensis"))

high_treatment <- ggplot(high_data, aes(x=event, y=Y_II_PAM, fill = species)) +
  geom_boxplot() +
  ylim(0,0.55) +
  scale_fill_manual(values = c("darkmagenta", "#E7298A","darkorange1")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  ggtitle("D) High") +
  xlab("Monitoring day") +
  ylab("YII") +
  theme(legend.position = "none") +
  scale_x_discrete(labels=c("Pre-stress", "1", "7")) +
  facet_wrap(vars(species), strip.position = "top", labeller = labeller(species = proper_names))
  

high_treatment

#mid
mid_data <- data_to_model %>%
  filter(treatment == "mid") %>%
  filter(event == "pre_stress" | event == "post_stress_1" | event == "post_stress_7") %>%
  mutate(species = fct_relevel(species, "cham", "calli ", "chillensis"))

mid_treatment <- ggplot(mid_data, aes(x=event, y=Y_II_PAM, fill = species)) +
  geom_boxplot() +
  scale_fill_manual(values = c("darkmagenta", "#E7298A","darkorange1")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  ylim(0,0.55) +
  ggtitle("C) Mid") +
  xlab("Monitoring day") +
  ylab("YII") +
  theme(legend.position = "none") +
  scale_x_discrete(labels=c("Pre-stress", "1", "7")) +
  facet_wrap(vars(species), strip.position = "top", labeller = labeller(species = proper_names))

mid_treatment

#low
low_data <- data_to_model %>%
  filter(treatment == "low") %>%
  filter(event == "pre_stress" | event == "post_stress_1" | event == "post_stress_7") %>%
  mutate(species = fct_relevel(species, "cham", "calli ", "chillensis"))

low_treatment <- ggplot(low_data, aes(x=event, y=Y_II_PAM, fill = species)) +
  geom_boxplot() +
  scale_fill_manual(values = c("darkmagenta", "#E7298A","darkorange1")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  ylim(0,0.55) +
  ggtitle("B) Low") +
  xlab("Monitoring day") +
  ylab("YII") +
  theme(legend.position = "none") +
  scale_x_discrete(labels=c("Pre-stress", "1", "7")) +
  facet_wrap(vars(species), strip.position = "top", labeller = labeller(species = proper_names))

low_treatment

#control
control_data <- data_to_model %>%
  filter(treatment == "control") %>%
  filter(event == "pre_stress" | event == "post_stress_1" | event == "post_stress_7") %>%
  mutate(species = fct_relevel(species, "cham", "calli ", "chillensis"))

control_treatment <- ggplot(control_data, aes(x=event, y=Y_II_PAM, fill = species)) +
  geom_boxplot() +
  scale_fill_manual(values = c("darkmagenta", "#E7298A","darkorange1")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  ylim(0,0.55) +
  ggtitle("A) Control") +
  xlab("Monitoring day") +
  ylab("YII") +
  theme(legend.position = "none") +
  scale_x_discrete(labels=c("Pre-stress", "1", "7")) +
  facet_wrap(vars(species), strip.position = "top", labeller = labeller(species = proper_names)) 


control_treatment

```

```{r}

stacked_plot <- (control_treatment/mid_treatment/low_treatment/high_treatment) 
  

stacked_plot
```

Want to test significance in change of YII for each treatment between days-
```{r}
#need a non-parametric comparison of multiple means, with paired data
#Found Friedman Test-

library(tidyverse)
library(rstatix)

#**for control per species**
#control chamberlanium
fried_cham_cont <- as.data.frame(control_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "cham")

fried_cham_cont_matrix <- as.matrix.data.frame(fried_cham_cont)

friedman.test(Y_II_PAM ~ event | id, fried_cham_cont_matrix)

#control calli
fried_calli_cont <- as.data.frame(control_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "calli ")

fried_calli_cont_matrix <- as.matrix.data.frame(fried_calli_cont)

friedman.test(Y_II_PAM ~ event | id, fried_calli_cont_matrix)

#control chile
fried_chile_cont <- as.data.frame(control_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "chillensis")

fried_chile_cont_matrix <- as.matrix.data.frame(fried_chile_cont)

friedman.test(Y_II_PAM ~ event | id, fried_chile_cont_matrix)

```

```{r}
#**for low per species**
#low chamberlanium
fried_cham_low <- as.data.frame(low_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "cham")

fried_cham_low_matrix <- as.matrix.data.frame(fried_cham_low)

friedman.test(Y_II_PAM ~ event | id, fried_cham_low_matrix)

#low calli
fried_calli_low <- as.data.frame(low_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "calli ")

fried_calli_low_matrix <- as.matrix.data.frame(fried_calli_low)

friedman.test(Y_II_PAM ~ event | id, fried_calli_low_matrix)

pairwise.wilcox.test(fried_calli_low$Y_II_PAM, g = fried_calli_low$event)

#low chile
fried_chile_low <- as.data.frame(low_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "chillensis")

fried_chile_low_matrix <- as.matrix.data.frame(fried_chile_low)

friedman.test(Y_II_PAM ~ event | id, fried_chile_low_matrix)

pairwise.wilcox.test(fried_chile_low$Y_II_PAM, g = fried_chile_low$event)
```

```{r}
#**for mid per species**
#mid chamberlanium
fried_cham_mid <- as.data.frame(mid_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "cham")

fried_cham_mid_matrix <- as.matrix.data.frame(fried_cham_mid)

friedman.test(Y_II_PAM ~ event | id, fried_cham_mid_matrix)

pairwise.wilcox.test(fried_cham_mid$Y_II_PAM, g = fried_cham_mid$event)

#mid calli
fried_calli_mid <- as.data.frame(mid_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "calli ")

fried_calli_mid_matrix <- as.matrix.data.frame(fried_calli_mid)

friedman.test(Y_II_PAM ~ event | id, fried_calli_mid_matrix)

pairwise.wilcox.test(fried_calli_mid$Y_II_PAM, g = fried_calli_mid$event)

#mid chile
fried_chile_mid <- as.data.frame(mid_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "chillensis")

fried_chile_mid_matrix <- as.matrix.data.frame(fried_chile_mid)

friedman.test(Y_II_PAM ~ event | id, fried_chile_mid_matrix)

pairwise.wilcox.test(fried_chile_mid$Y_II_PAM, g = fried_chile_mid$event)
```

```{r}
#**for high per species**
#high chamberlanium
fried_cham_high <- as.data.frame(high_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "cham")

fried_cham_high_matrix <- as.matrix.data.frame(fried_cham_high)

friedman.test(Y_II_PAM ~ event | id, fried_cham_high_matrix)

pairwise.wilcox.test(fried_cham_high$Y_II_PAM, g = fried_cham_high$event)

#high calli
fried_calli_high <- as.data.frame(high_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "calli ")

fried_calli_high_matrix <- as.matrix.data.frame(fried_calli_high)

friedman.test(Y_II_PAM ~ event | id, fried_calli_high_matrix)

pairwise.wilcox.test(fried_calli_high$Y_II_PAM, g = fried_calli_high$event)

#high chile
fried_chile_high <- as.data.frame(high_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "chillensis")

fried_chile_high_matrix <- as.matrix.data.frame(fried_chile_high)

friedman.test(Y_II_PAM ~ event | id, fried_chile_high_matrix)

pairwise.wilcox.test(fried_chile_high$Y_II_PAM, g = fried_chile_high$event)
```







Code for color species comparison plot_
```{r}
color_high <- colour_data %>%
  filter(treatment == "high")

color_high_plot<- ggplot(high_data, aes(x= species, y= deltaint_mean ,fill = species)) +
  geom_boxplot()
  
color_high_plot
```








BELOW THIS IS JUST WORKING OUT STUFF

filtering colour data:
```{r}
colour_data <- data %>%
  unite(id, species, species_card, sep = '_', remove = F) %>%
  dplyr::select("date","yellow_tub","species_card","species","number","treatment","name","event","int_mean","id") %>%
  mutate(id = as.factor(id)) %>% 
  filter(event == "pre_stress" | event == "post_stress_7")

colour_data <- colour_data %>% mutate(event = fct_relevel(event, "pre_stress", "post_stress_7")) %>%
  na.omit()
```
#####WANT to make this change in intensity over time



plot it:high_
```{r}
high_data <- colour_data %>%
  filter(treatment == "high")

high_plot<- ggplot(high_data, aes(x= event, y= int_mean ,fill = species)) +
  geom_boxplot()
  
high_plot
```

control_
```{r}
control_data <- colour_data %>%
  filter(treatment == "control")

control_plot<- ggplot(control_data, aes(x= event, y= int_mean ,fill = species)) +
  geom_boxplot()
  
control_plot
```



