---
title: "Results Stat Analysis"
author: "Lucy and Sammie"
date: "25/11/2021"
output: pdf_document
---

Load libraries:
```{r, include=FALSE}
library(ggplot2)
library(popbio)
library(dplyr)
library(patchwork)
library(DHARMa)
library(mgcv)
library(fitdistrplus)
library(goft)
library(gamlss)
library(viridis)
library(tidyverse)
```

Load in data and rename PAM column:
```{r}
data <- read.csv("test_data_NOV25.csv")
data <- data %>%
rename("Y_II_PAM" = "Y_II..PAM.") 
```

Setting up data with PAM columns only:
```{r}
data_to_model <- data %>%
  unite(id, species, species_card, sep = '_', remove = F) %>%
  dplyr::select("date","yellow_tub","species_card","species","number","treatment","name","event","Y_II_PAM","id") %>%
  mutate(id = as.factor(id))
```

re-ordering event for graphing later:
```{r}
data_to_model <- data_to_model %>% mutate(event = fct_relevel(event, "pre_stress", "post_stress", "post_stress_1", "post_stress_2", "post_stress_3", "post_stress_4", "post_stress_5", "post_stress_6","post_stress_7"))
```

testing for normality:
```{r}
shapiro.test(data_to_model$Y_II_PAM)

#not normal P=3.09e-08
```

Cullen & Frey graph:
```{r}
descdist(data_to_model$Y_II_PAM)
```
Looks like it might be a gamma/lognormal/weibull

Look at histograms to find distribution:
```{r}
mGA <- histDist(data_to_model$Y_II_PAM, "GA", density = T, main = "Gamma")
mlN <- histDist(data_to_model$Y_II_PAM, "LOGNO", density = T, main = "lNorm")
mWE <- histDist(data_to_model$Y_II_PAM, "WEI", density = T, main = "Weibull")
mNO <- histDist(data_to_model$Y_II_PAM, "NO", density = T, main = "Normal")
```
```{r}
GAIC(mGA, mlN, mWE, mNO, mSEP2)
#these give us negative values, which fit is best????
```


```{r}
fitDist(Y_II_PAM, data = data_to_model, type = "realAll", try.gamlss = T)
```

Trying skewed exponential power type 2:
```{r}
mSEP2 <- histDist(data_to_model$Y_II_PAM, "SEP2", density = T, main = "SEP2")
```
Decided to move forward exploring SEP2 and WEI distributions:
```{r}
mod_SEP2 <- gamlssML(Y_II_PAM ~ species + treatment*event + random(id), family = SEP2(), data = data_to_model, method = nlminb())

summary(mod_SEP2)
```
Model using WEI-
```{r}
mod_WEI <- gamlss(Y_II_PAM ~ species + treatment*event + random(id), family = WEI(), data = data_to_model, method = RS(), control = gamlss.control())

summary(mod_WEI)
```

Line-range plot--
```{r}

```

Sammie's test of colour dataframe and cleaning
```{r}
write.csv(data_to_model,"id_data.csv")


c_data <- read.csv("id_edit.csv") %>%
  filter(event == "pre_stress" | event == "post_stress_7") %>%
   mutate(id = as.factor(id))

ct_data <- read.csv("test_colour.csv") %>%
  mutate(id = as.factor(id)) %>%
  filter(yellow_tub == "1"| yellow_tub == "3") %>%
  dplyr::select(id, change_g7, change_r7, species, treatment) %>%
  na.omit(id, change_g7, change_r7, species, treatment)
  
```

filter for species
```{r}
cham_ctdata <- ct_data %>%
  mutate(treatment = fct_relevel(treatment, "control", "low", "mid", "high")) %>%
  filter(species == "cham")

ca_ctdata <- ct_data %>%
  mutate(treatment = fct_relevel(treatment, "control", "low", "mid", "high")) %>%
  filter(species == "calli")

chil_ctdata <- ct_data %>%
  mutate(treatment = fct_relevel(treatment, "control", "low", "mid", "high")) %>%
  filter(species == "chillensis")
```


Normality test for colour data
```{r}

shapiro.test(c_data$change_grey_f)

shapiro.test(c_data$change_red_f)
```


Collen & Frey Graph of colour data
```{r}
descdist(c_data$change_grey_f)

descdist(c_data$change_red_f)

```

Looking at other distribution fits
```{r}
fitDist(Y_II_PAM, data = data_to_model, type = "realAll", try.gamlss = T)

```

Look at histograms to find distribution: *insert depending on results of descdist
```{r}
mGA <- histDist(data_to_model$Y_II_PAM, "GA", density = T, main = "Gamma")
mlN <- histDist(data_to_model$Y_II_PAM, "LOGNO", density = T, main = "lNorm")
mWE <- histDist(data_to_model$Y_II_PAM, "WEI", density = T, main = "Weibull")
mNO <- histDist(data_to_model$Y_II_PAM, "NO", density = T, main = "Normal")
```

AIC 
```{r}
GAIC(mGA, mlN, mWE, mNO, mSEP2)
```

Testing out the lowest AIC 

```{r}
mSEP2 <- histDist(data_to_model$Y_II_PAM, "SEP2", density = T, main = "SEP2")
```

Boxplot for colour data
```{r}
cham_ctgplot <- ggplot(cham_ctdata, aes(x= treatment, y = change_g7, fill = treatment)) +
  geom_boxplot() + 
  ylab("Change in grey scale brightness") +
  xlab("Treatment") +
  ggtitle("Chamberlainium") +
  theme_bw()

cham_ctrplot <- ggplot(cham_ctdata, aes(x= treatment, y = change_r7, fill = treatment)) +
  geom_boxplot() + 
  ylab("Change in red scale brightness") +
  xlab("Treatment") +
  ggtitle("Chamberlainium") +
  theme_bw()

cham_ctgplot
cham_ctrplot

ca_ctgplot <- ggplot(ca_ctdata, aes(x= treatment, y = change_g7, fill = treatment)) +
  geom_boxplot() + 
  ylab("Change in grey scale brightness") +
  xlab("Treatment") +
  ggtitle("Calliarthron") +
  theme_bw()

ca_ctrplot <- ggplot(ca_ctdata, aes(x= treatment, y = change_r7, fill = treatment)) +
  geom_boxplot() + 
  ylab("Change in red scale brightness") +
  xlab("Treatment") +
  ggtitle("Calliarthron") +
  theme_bw()

ca_ctgplot
ca_ctrplot

chil_ctgplot <- ggplot(chil_ctdata, aes(x= treatment, y = change_g7, fill = treatment)) +
  geom_boxplot() + 
  ylab("Change in grey scale brightness") +
  xlab("Treatment") +
  ggtitle("Chilensis") +
  theme_bw()

chil_ctrplot <- ggplot(chil_ctdata, aes(x= treatment, y = change_r7, fill = treatment)) +
  geom_boxplot() + 
  ylab("Change in red scale brightness") +
  xlab("Treatment") +
  ggtitle("Chilensis") +
  theme_bw()

chil_ctgplot
chil_ctrplot

```

