---
title: "Results Stat Analysis"
author: "Lucy"
date: "25/11/2021"
output: pdf_document
---

Load libraries:
```{r, include=FALSE}
library(ggplot2)
library(popbio)
library(dplyr)
library(patchwork)
library(DHARMa)
library(mgcv)
library(fitdistrplus)
library(goft)
library(gamlss)
library(viridis)
library(tidyverse)
library(rstatix)
```

Load in data and rename PAM column: 
```{r}
data <- read.csv("PAM_data_DS2021.csv")
data <- data %>%
rename("Y_II_PAM" = "Y_II..PAM.") %>%
  mutate(treatment = fct_relevel(treatment, "control", "low", "mid", "high")) %>%
  mutate(species = fct_relevel(species, "cham", "calli", "chillensis"))
```

Setting up data with PAM-data columns only:
```{r}
data_to_model <- data %>%
  unite(id, species, species_card, sep = '_', remove = F) %>%
  dplyr::select("date","yellow_tub","species_card","species","number","treatment","name","event","Y_II_PAM","id") %>%
  mutate(id = as.factor(id))
```

Re-ordering "event" column for graphing later:
```{r}
data_to_model <- data_to_model %>% mutate(event = fct_relevel(event, "pre_stress", "post_stress", "post_stress_1", "post_stress_2", "post_stress_3", "post_stress_4", "post_stress_5", "post_stress_6","post_stress_7"))
```

Testing YII data for normality:
```{r}
shapiro.test(data_to_model$Y_II_PAM)

#not normal P=3.09e-08
```

Making Cullen & Frey graph:
```{r}
descdist(data_to_model$Y_II_PAM)
```
Looks like it might be a gamma/lognormal/weibull

fitDist to get recommended distribution:
```{r}
fitDist(Y_II_PAM, data = data_to_model, type = "realAll", try.gamlss = T)
```

Look at histograms to find distribution fit:
```{r}
mGA <- histDist(data_to_model$Y_II_PAM, "GA", density = T, main = "Gamma")
mlN <- histDist(data_to_model$Y_II_PAM, "LOGNO", density = T, main = "lNorm")
mWE <- histDist(data_to_model$Y_II_PAM, "WEI", density = T, main = "Weibull")
mNO <- histDist(data_to_model$Y_II_PAM, "NO", density = T, main = "Normal")
mSEP2 <- histDist(data_to_model$Y_II_PAM, "SEP2", density = T, main = "SEP2")
```
```{r}
GAIC(mGA, mlN, mWE, mNO, mSEP2)
#WEI is really close to SEP2 but has fewer degrees of freedom (we chose to move forward with WEI)
```


Full model for all event days using WEI:
```{r}
mod_WEI <- gamlss(Y_II_PAM ~ species + treatment*event + random(id), family = WEI(), data = data_to_model, method = RS(), control = gamlss.control())

summary(mod_WEI)
```

Modelling simple version -just 3 event days: pre-stress, post_stress1, and post_stress7:
```{r}
simpler_data <- data_to_model %>%
  filter(event == "pre_stress" | event == "post_stress_1" | event == "post_stress_7")

simpler_mod_WEI <- gamlss(Y_II_PAM ~ species + treatment*event + random(id), family = WEI(), data = simpler_data, method = RS(), control = gamlss.control())

summary(simpler_mod_WEI)
```


setting up plotting data and making boxplot to visualize data for chamberlanium:
```{r}
ch_f <- filter(data_to_model, species == "cham") %>%
  na.omit()

ch_f$event = as.factor(ch_f$event)

ch_f <- ch_f %>% mutate(treatment = fct_relevel(treatment, "control", "low", "mid", "high"))

head(ch_f)

cham_plot<- ggplot(ch_f, aes(x= event, y= Y_II_PAM,fill = treatment)) +
  geom_boxplot()
  
cham_plot
```
setting up plotting data and boxplot to visualize calliarthron data:
```{r}
calli_f <- filter(data_to_model, species == "calli") %>%
  na.omit()

calli_f$event = as.factor(calli_f$event)

calli_f <- calli_f %>% mutate(treatment = fct_relevel(treatment, "control", "low", "mid", "high"))

head(calli_f)

calli_plot<- ggplot(calli_f, aes(x= event, y= Y_II_PAM,fill = treatment)) +
  geom_boxplot()
  
calli_plot
```
setting up plotting data and boxplot for chilensis data:
```{r}
chil_f <- filter(data_to_model, species == "chillensis") %>%
  na.omit()

chil_f$event = as.factor(chil_f$event)

chil_f <- chil_f %>% mutate(treatment = fct_relevel(treatment, "control", "low", "mid", "high"))

head(chil_f)

chil_plot<- ggplot(chil_f, aes(x= event, y= Y_II_PAM,fill = treatment)) +
  geom_boxplot()
  
chil_plot
```

Making a box+whisker plot for each of the three species separately, following YII over observation week:
```{r}
library(plotrix)
library(tidyverse)

#Chamberlanium
ch_simple <- filter(ch_f, event == "pre_stress" | event == "post_stress_1" | event == "post_stress_2" | event == "post_stress_3" | event == "post_stress_4" | event == "post_stress_5" | event == "post_stress_6" | event == "post_stress_7")

ch_for_plot <- ch_simple %>%
group_by(treatment, event) %>%
mutate(mean = mean(Y_II_PAM), SE = std.error(Y_II_PAM))
  

chamberlanium_plot <- ggplot(data = ch_for_plot, aes(x = event, y = mean, group = treatment, color = treatment)) +
  geom_point(aes(shape=treatment, color=treatment), size = 3) +
  ylim(0,0.4) +
  scale_color_manual(values=c('#3C507BFF', '#C7E993', '#DC7B00', '#B11901FF')) +
  scale_shape_manual(values=c(1,16,17,15)) +
  scale_fill_discrete(name = "H202 Concentration", labels = c("Control", "Low", "Mid", "High")) +
  ggtitle("A) C. tumidum") +
  xlab("") +
  ylab("YII") +
  geom_linerange(aes(ymin = mean - SE, ymax = mean + SE)) +
  geom_line() +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 25), axis.title.y = element_text(size = 25), legend.title = element_text(size = 20), axis.text.y = element_text(size = 20), title = element_text(size = 20)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(legend.position = "none") +
  scale_x_discrete(labels=c("Pre-stress", "1", "2", "3", "4", "5", "6", "7")) +
  geom_vline(xintercept = 1.5, linetype = "dashed", colour ="grey") 
  

chamberlanium_plot

```
plot for calliarthron:
```{r}
#calliarthron
calli_simple <- filter(calli_f, event == "pre_stress" | event == "post_stress_1" | event == "post_stress_2" | event == "post_stress_3" | event == "post_stress_4" | event == "post_stress_5" | event == "post_stress_6" | event == "post_stress_7")

calli_for_plot <- calli_simple %>%
group_by(treatment, event) %>%
mutate(mean = mean(Y_II_PAM), SE = std.error(Y_II_PAM))
  

calliarthron_plot <- ggplot(data = calli_for_plot, aes(x = event, y = mean, group = treatment, color = treatment)) +
  geom_point(aes(shape=treatment, color=treatment), size = 3) +
  ylim(0,0.4) +
  scale_color_manual(values=c('#3C507BFF', '#C7E993', '#DC7B00', '#B11901FF')) +
  scale_shape_manual(values=c(1,16,17,15)) +
  scale_fill_discrete(name = "H202 Concentration", labels = c("Control", "Low", "Mid", "High")) +
  ggtitle("B) C. tuberculosum") +
  xlab("Monitoring Day") +
  ylab("YII") +
  geom_linerange(aes(ymin = mean - SE, ymax = mean + SE)) +
  geom_line() +
  theme_classic() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 25), axis.title.y = element_text(size = 25), legend.title = element_text(size = 20), axis.text.y = element_text(size = 20), title = element_text(size = 20)) +
  scale_x_discrete(labels=c("Pre-stress", "1", "2", "3", "4", "5", "6", "7")) +
  geom_vline(xintercept = 1.5, linetype = "dashed", colour ="grey")   

calliarthron_plot
```
plot for chilensis:
```{r}
#Chilensis
chil_simple <- filter(chil_f, event == "pre_stress" | event == "post_stress_1" | event == "post_stress_2" | event == "post_stress_3" | event == "post_stress_4" | event == "post_stress_5" | event == "post_stress_6" | event == "post_stress_7")

chil_for_plot <- chil_simple %>%
group_by(treatment, event) %>%
mutate(mean = mean(Y_II_PAM), SE = std.error(Y_II_PAM))
  

chilensis_plot <- ggplot(data = chil_for_plot, aes(x = event, y = mean, group = treatment, color = treatment)) +
  geom_point(aes(shape=treatment, color=treatment), size = 3) +
  ylim(0,0.4) +
  scale_color_manual(values=c('#3C507BFF', '#C7E993', '#DC7B00', '#B11901FF')) +
  scale_shape_manual(values=c(1,16,17,15)) +
  scale_fill_discrete(name = "H202 Concentration", labels = c("Control", "Low", "Mid", "High")) +
  ggtitle("C) C. chilensis") +
  xlab("") +
  ylab("YII") +
  geom_linerange(aes(ymin = mean - SE, ymax = mean + SE)) +
  geom_line() +
  theme_classic() +
  theme(legend.title = "Treatment", legend.key.size = unit(2,"cm"), legend.text = element_text(size= 16), axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 30), axis.title.y = element_text(size = 25), legend.title = element_text(size = 20), axis.text.y = element_text(size = 20), title = element_text(size = 20)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  scale_x_discrete(labels=c("Pre-stress", "1", "2", "3", "4", "5", "6", "7")) +
  geom_vline(xintercept = 1.5, linetype = "dashed", colour ="grey") 

chilensis_plot
```

Patchworking the three above plots together:
```{r}

chamberlanium_plot + calliarthron_plot + chilensis_plot

pam_plots_fig1 <- chamberlanium_plot + calliarthron_plot + chilensis_plot

ggsave(pam_plots_fig1, filename = "pam_fig1.png", height = 10, width = 20)
```
Making boxplots for each treatment, comparing species:
```{r}

proper_names <- c(cham = "C. tumidum", "calli" = "C. tuberculosum", chillensis = "C. chilensis")

#high
high_data <- data_to_model %>%
  filter(treatment == "high") %>%
  filter(event == "pre_stress" | event == "post_stress_1" | event == "post_stress_7") %>%
  mutate(species = fct_relevel(species, "cham", "calli", "chillensis"))

high_treatment <- ggplot(high_data, aes(x=event, y=Y_II_PAM, fill = species)) +
  geom_boxplot() +
  ylim(0,0.55) +
  scale_fill_manual(values = c("#4E63AA", "#DD6E96","#FCDA61")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(strip.text.x = element_text(size = 20, face = "italic"), axis.text.x = element_text(size = 20), axis.title.x = element_text(size = 30), axis.title.y = element_text(size = 30), legend.title = element_text(size = 20), axis.text.y = element_text(size = 20), title = element_text(size = 30)) +
  ggtitle("D) High") +
  xlab("Monitoring day") +
  ylab("YII") +
  theme(legend.position = "none") +
  scale_x_discrete(labels=c("Pre-stress", "1", "7")) +
  facet_wrap(vars(species), strip.position = "top", labeller = labeller(species = proper_names)) +
  geom_vline(xintercept = 1.5, linetype = "dashed", colour ="grey")
  

high_treatment

#mid
mid_data <- data_to_model %>%
  filter(treatment == "mid") %>%
  filter(event == "pre_stress" | event == "post_stress_1" | event == "post_stress_7") %>%
  mutate(species = fct_relevel(species, "cham", "calli", "chillensis"))

mid_treatment <- ggplot(mid_data, aes(x=event, y=Y_II_PAM, fill = species)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#4E63AA", "#DD6E96","#FCDA61")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme (strip.text.x = element_text(size = 20, face = "italic"), axis.text.x = element_text(size = 20), axis.title.x = element_text(size = 30), axis.title.y = element_text(size = 30), legend.title = element_text(size = 20), axis.text.y = element_text(size = 20), title = element_text(size = 30)) +
  ylim(0,0.55) +
  ggtitle("C) Mid") +
  xlab("") +
  ylab("YII") +
  theme(legend.position = "none") +
  scale_x_discrete(labels=c("Pre-stress", "1", "7")) +
  facet_wrap(vars(species), strip.position = "top", labeller = labeller(species = proper_names)) +
  geom_vline(xintercept = 1.5, linetype = "dashed", colour ="grey")

mid_treatment

#low
low_data <- data_to_model %>%
  filter(treatment == "low") %>%
  filter(event == "pre_stress" | event == "post_stress_1" | event == "post_stress_7") %>%
  mutate(species = fct_relevel(species, "cham", "calli", "chillensis"))

low_treatment <- ggplot(low_data, aes(x=event, y=Y_II_PAM, fill = species)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#4E63AA", "#DD6E96","#FCDA61")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme (strip.text.x = element_text(size = 20, face = "italic"), axis.text.x = element_text(size = 20), axis.title.x = element_text(size = 30), axis.title.y = element_text(size = 30), legend.title = element_text(size = 20), axis.text.y = element_text(size = 20), title = element_text(size = 30)) +
  ylim(0,0.55) +
  ggtitle("B) Low") +
  xlab("") +
  ylab("YII") +
  theme(legend.position = "none") +
  scale_x_discrete(labels=c("Pre-stress", "1", "7")) +
  facet_wrap(vars(species), strip.position = "top", labeller = labeller(species = proper_names)) +
  geom_vline(xintercept = 1.5, linetype = "dashed", colour ="grey")

low_treatment

#control
control_data <- data_to_model %>%
  filter(treatment == "control") %>%
  filter(event == "pre_stress" | event == "post_stress_1" | event == "post_stress_7") %>%
  mutate(species = fct_relevel(species, "cham", "calli", "chillensis"))

control_treatment <- ggplot(control_data, aes(x=event, y=Y_II_PAM, fill = species)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#4E63AA", "#DD6E96","#FCDA61")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme (strip.text.x = element_text(size = 20, face = "italic"), axis.text.x = element_text(size = 20), axis.title.x = element_text(size = 30), axis.title.y = element_text(size = 30), legend.title = element_text(size = 20), axis.text.y = element_text(size = 20), title = element_text(size = 30)) +
  ylim(0,0.55) +
  ggtitle("A) Control") +
  xlab("") +
  ylab("YII") +
  theme(legend.position = "none") +
  scale_x_discrete(labels=c("Pre-stress", "1", "7")) +
  facet_wrap(vars(species), strip.position = "top", labeller = labeller(species = proper_names)) +
  geom_vline(xintercept = 1.5, linetype = "dashed", colour ="grey")


control_treatment

ggsave(control_treatment, filename = "pam_control_fig2.png", height = 10, width = 15)
ggsave(low_treatment, filename = "pam_low_fig2.png", height = 10, width = 15)
ggsave(mid_treatment, filename = "pam_mid_fig2.png", height = 10, width = 15)
ggsave(high_treatment, filename = "pam_high_fig2.png", height = 10, width = 15)

```

Testing significance in change of YII for each treatment, between days. separated below by treatment and species, with corresponding Friedman and Wilcoxon tests to test significance:
```{r}
#need a non-parametric comparison of multiple means, with paired data

#Friedman Test- followed by Wilcoxon paired comparison of means

library(tidyverse)
library(rstatix)

#**for control per species**

#filter datasets for control treatment and each individual species, and run comparison of means stats tests

#control chamberlanium: 
fried_cham_cont <- as.data.frame(control_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "cham")

fried_cham_cont_matrix <- as.matrix.data.frame(fried_cham_cont)

friedman.test(Y_II_PAM ~ event | id, fried_cham_cont_matrix)

#control calli
fried_calli_cont <- as.data.frame(control_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "calli")

fried_calli_cont_matrix <- as.matrix.data.frame(fried_calli_cont)

friedman.test(Y_II_PAM ~ event | id, fried_calli_cont_matrix)

#control chile
fried_chile_cont <- as.data.frame(control_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "chillensis")

fried_chile_cont_matrix <- as.matrix.data.frame(fried_chile_cont)

friedman.test(Y_II_PAM ~ event | id, fried_chile_cont_matrix)

```

```{r}
#**for low treatment per species**

#filter datasets for low treatment and each individual species, and run comparison of means stats tests

#low chamberlanium
fried_cham_low <- as.data.frame(low_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "cham")

fried_cham_low_matrix <- as.matrix.data.frame(fried_cham_low)

friedman.test(Y_II_PAM ~ event | id, fried_cham_low_matrix)

#low calli
fried_calli_low <- as.data.frame(low_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "calli")

fried_calli_low_matrix <- as.matrix.data.frame(fried_calli_low)

friedman.test(Y_II_PAM ~ event | id, fried_calli_low_matrix)

pairwise.wilcox.test(fried_calli_low$Y_II_PAM, g = fried_calli_low$event)

#low chile
fried_chile_low <- as.data.frame(low_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "chillensis")

fried_chile_low_matrix <- as.matrix.data.frame(fried_chile_low)

friedman.test(Y_II_PAM ~ event | id, fried_chile_low_matrix)

pairwise.wilcox.test(fried_chile_low$Y_II_PAM, g = fried_chile_low$event)
```

```{r}
#**for mid treatment per species**

#filter datasets for mid treatment and each individual species, and run comparison of means stats tests

#mid chamberlanium
fried_cham_mid <- as.data.frame(mid_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "cham")

fried_cham_mid_matrix <- as.matrix.data.frame(fried_cham_mid)

friedman.test(Y_II_PAM ~ event | id, fried_cham_mid_matrix)

pairwise.wilcox.test(fried_cham_mid$Y_II_PAM, g = fried_cham_mid$event)

#mid calli
fried_calli_mid <- as.data.frame(mid_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "calli")

fried_calli_mid_matrix <- as.matrix.data.frame(fried_calli_mid)

friedman.test(Y_II_PAM ~ event | id, fried_calli_mid_matrix)

pairwise.wilcox.test(fried_calli_mid$Y_II_PAM, g = fried_calli_mid$event)

#mid chile
fried_chile_mid <- as.data.frame(mid_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "chillensis")

fried_chile_mid_matrix <- as.matrix.data.frame(fried_chile_mid)

friedman.test(Y_II_PAM ~ event | id, fried_chile_mid_matrix)

pairwise.wilcox.test(fried_chile_mid$Y_II_PAM, g = fried_chile_mid$event)
```

```{r}
#**for high treatment per species**

#filter datasets for high treatment and each individual species, and run comparison of means stats tests

#high chamberlanium
fried_cham_high <- as.data.frame(high_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "cham")

fried_cham_high_matrix <- as.matrix.data.frame(fried_cham_high)

friedman.test(Y_II_PAM ~ event | id, fried_cham_high_matrix)

pairwise.wilcox.test(fried_cham_high$Y_II_PAM, g = fried_cham_high$event)

#high calli
fried_calli_high <- as.data.frame(high_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "calli")

fried_calli_high_matrix <- as.matrix.data.frame(fried_calli_high)

friedman.test(Y_II_PAM ~ event | id, fried_calli_high_matrix)

pairwise.wilcox.test(fried_calli_high$Y_II_PAM, g = fried_calli_high$event)

#high chile
fried_chile_high <- as.data.frame(high_data) %>%
  select(c("event", "Y_II_PAM", "id", "species")) %>%
  filter(species == "chillensis")

fried_chile_high_matrix <- as.matrix.data.frame(fried_chile_high)

friedman.test(Y_II_PAM ~ event | id, fried_chile_high_matrix)

pairwise.wilcox.test(fried_chile_high$Y_II_PAM, g = fried_chile_high$event)
```






